---
title: "Reanalysis of 'Benefit valuation predicts gratitude'"
author: "Forster, Pedersen, Smith, McCullough, & Lieberman"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning = F, message = F}
library(tidyverse)
library(lmerTest)
library(semTools)
```

# About this Reanalysis

These are re-analyses (and alternative analyses) to those presented in 'Benefit valuation predicts gratitude' (Forster et al., 2017).

See the companion syntax file for data cleaning procedures that produced `grat_wtr_study1_clean.csv`.

# Study 1

```{r message = F}
study1_clean <- read_csv("grat_wtr_study1_clean.csv")
```


## Boxplots

WTR other->self

```{r}
study1_clean %>%
  # just to get one WTR estimate per person, remove all but one row/person
  filter(scenario == "S1") %>% 
  ggplot(aes(x = cond, y = wtr_os_est)) +
  geom_boxplot() +
  labs(title = "Study 1 WTR_os by relationship type", x = "Relationship Type", y = "WTR_os") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = .8))
```

WTR self->other

```{r}
study1_clean %>%
  # just to get one WTR estimate per person, remove all but one row/person
  filter(scenario == "S1") %>% 
  ggplot(aes(x = cond, y = wtr_so_est)) +
  geom_boxplot() +
  labs(title = "Study 1 WTR_so by relationship type", x = "Relationship Type", y = "WTR_so") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = .8))
```


## Analyses


Gratitude as outcome

```{r}
grat_out <- lmer(grateful ~ wtr_os_est * mean_log_val + (1 | ResponseId), data = study1_clean)
summary(grat_out)
```

Obligation as outcome

```{r}
oblig_out <- lmer(obligated ~ wtr_os_est * mean_log_val + (1 + mean_log_val | ResponseId), data = study1_clean)
summary(oblig_out)
```

Want to Return as outcome

```{r}
want_out <- lmer(want ~ wtr_os_est * mean_log_val + (1 + mean_log_val | ResponseId), data = study1_clean)
summary(want_out)
```

Happiness as outcome

```{r}
happy_out <- lmer(happy ~ wtr_os_est * mean_log_val + (1 + mean_log_val | ResponseId), data = study1_clean)
summary(happy_out)
```


# Study 2

```{r message = F}
study2_clean <- read_csv("grat_wtr_study2_clean.csv")
```

## Boxplots

WTR other->self

```{r}
study2_clean %>% 
  ggplot(aes(x = cond, y = wtr_os_est)) +
  geom_boxplot() +
  labs(title = "Study 2 WTR_os across Relationship Type", x = "Relationship Type", y = "WTR_os") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = .8))
```


WTR self->other

```{r}
study2_clean %>% 
  ggplot(aes(x = cond, y = wtr_so_est)) +
  geom_boxplot() +
  labs(title = "Study 2 WTR_so across Relationship Type", x = "Relationship Type", y = "WTR_so") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = .8))
```

## Analyses


Mean-center WTR variables

```{r}
study2_clean <- study2_clean %>% 
  mutate(wtr_so_mc = wtr_so_est - mean(wtr_so_est, na.rm = T),
         wtr_os_mc = wtr_os_est - mean(wtr_os_est, na.rm = T),
         log_ben_mc = log_ben - mean(log_ben, na.rm = T),
         wtr_so_int = wtr_so_mc * log_ben_mc,
         wtr_os_int = wtr_os_mc * log_ben_mc)
```



```{r}

grat_model <- '
  gratitude =~ NA*grateful + thankful + appreciative
  gratitude ~~ 1*gratitude
  gratitude ~ 0*1
  
  gratitude ~ wtr_os_mc + log_ben_mc + wtr_os_int
'

grat_sem <- sem(grat_model, data = study2_clean, ordered = c("grateful", "thankful", "appreciative"))

summary(grat_sem, standardized = T, rsquare = T)

```

In this model, only about .5% of variance is attributable to WTR_os

```{r}
grat_so_model <- '
  gratitude =~ NA*grateful + thankful + appreciative
  gratitude ~~ 1*gratitude
  gratitude ~ 0*1
  
  gratitude ~ wtr_so_mc + log_ben_mc + wtr_so_int
'

grat_so_sem <- sem(grat_so_model, data = study2_clean, ordered = c("grateful", "thankful", "appreciative"))

summary(grat_so_sem, standardized = T, rsquare = T)
```

Here, WTR_so only accounts for about 1.3% of variance in gratitude.

```{r}
indebt_os_model <- '
  indebt =~ NA*indebted + burdened + required + owed
  indebt ~~ 1*indebt
  indebt ~ 0*1
  
  indebt ~ wtr_os_mc + log_ben_mc + wtr_os_int
'

indebt_os_sem <- sem(indebt_os_model, data = study2_clean, ordered = c("indebted","burdened", "required", "owed"))

summary(indebt_os_sem, standardized = T, rsquare = T)
```

```{r}
indebt_so_model <- '
  indebt =~ NA*indebted + burdened + required + owed
  indebt ~~ 1*indebt
  indebt ~ 0*1
  
  indebt ~ wtr_so_mc + log_ben_mc + wtr_so_int
'

indebt_so_sem <- sem(indebt_so_model, data = study2_clean, ordered = c("indebted","burdened", "required", "owed"))

summary(indebt_so_sem, standardized = T, rsquare = T)
```
